Tempus: Probabilistic Network Latency Verifier

### 2022-12-02 paper/main.tex
[Paper] added explanation for both runtime and opt

Spec
\begin{itemize}
    \item Julia
    \item i7-1065G7 16GB RAM
\end{itemize}

Network model
\begin{itemize}
    \item OSPF weight 1
    \item 0.1\% link failure rate
    \item $10^{-8}$ inaccuracy
    \item Reachability between two edge nodes
    \item Queuing latency: LogNormal distribution
    \item Link latency: either Gamma or Chisq (chosen at random)
    \item Bounded reachability with t=1000.0
\end{itemize}

@@

\begin{itemize}
    \item The convolution step takes the most amount of time
    \item \textbf{Runtime performance correlates with the amount of convolutions} $->$ plot
    \item The amount of convolution is the result of the optimizations that we've done
\end{itemize}

@@

\begin{itemize}
    \item While the amount of path in those states can explain the amount of EC reduction, to get to 
        the amount of convolution, we also need to consider the length of each of those path
    \item By considering the both amount of path and its length to counting the convolution, we could 
        compare consolidation and memoization strategy to the unoptimized version
    \item We could see that in the best case, the optimization strategies \textbf{reduce the amount of 
        convolution by 98\%} (in the case of GtsCe)
\end{itemize}

@@

\section{Related Works}

>>>>

To evaluate our idea, we implemented Tempus in $\sim600$ lines of Julia code.
This code will take a topology configuration, source and destination node, and the 
temporal property in question and its threshold.
It will then output the final probability of the network upholding that temporal property.
We ran our prototype on a machine with i7-1065G7 CPU and 16 GB of RAM.

We tested Tempus on 5 real-world WAN topologies from the Topology Zoo and 3 datacenter topologies 
(fat-tree, with various amount of pods).
We use OSPF as the routing protocol with uniform link weights of 1 and 0.1\% failure rate in all 
links, in accordance to the failure rate in previous studies. %TODO: cite
We then run the functional verification with $10^{-8}$ inaccuracy level.

To evaluate the performance and scalability of Tempus, we first evaluate the running time of the 
temporal verification on different topologies given a mixture of latency distributions. 
We then examine the effectiveness of our optimization technique by comparing the number of 
convolution of the former results with the unoptimized version.
Next, we compare the result of the consolidation strategy in terms of the number of equivalence 
classes, instead of convolution.
Finally, we compare the effect of latency distribution type on the performance of Tempus.

Since the CDF operation that we define for temporal property verification is a relatively cheap 
operation, we expect our result to generalize with any threshold we chose. 
In other words, while changing the threshold will change the final probability, it won't change 
the runtime performance of Tempus.
Thus, we set an arbitrary threshold for our evaluation.

@@

To begin our evaluation, we first define the $l_p$ and $l_q$ for each component in 
the topology with varying probability distributions to ensure the robustness of our method.
Some distributions (like LogNormal) don't have a closed-form solution for their convolution, even 
against itself, while others (e.g. Gamma, Chi-Squared) do.
Since our algorithm relies on convolution, having a diversity of distributions, especially accross
these categories, is important for the representativeness of our evaluation.
In light of this, we set all the queuing latency $l_q$ of our topologies' component to be LogNormal 
and the link latency $l_p$ to be randomly chosen between Gamma and Chi-Squared.

We measure the runtime performance of Tempus by verifying the bounded reachability property of two 
edge routers.
Fig. \ref{fig:scalability} shows the temporal verification runtime performance (excluding functional 
verification) and the amount of convolution operation performed in each network.
We note that we can precisely verify the temporal property of networks with hundreds of links in 
the order of minutes.

The convolution step of Tempus takes the most amount of time in the temporal verification process, 
hence we can see in the same figure that the runtime performance of Tempus and the total amount 
of convolution is highly correlated.

We conclude that \textbf{the performance of Tempus is primarily bottlenecked by the convolution 
operation} and for a reasonably low imprecision level, \textbf{Tempus runs in the order of 
minutes}.

We have established that the convolution operation is a relatively expensive operation within 
Tempus.
Now, we inspect the effectiveness of our optimization techniques in reducing them.

Fig. \ref{fig:opt} shows the effect of the two optimization techniques, consolidation and 
memoization, in reducing the amount of convolution (in logarithmic scale).
We exclude the third optimization technique (grouping) in this evaluation because they are not 
designed to reduce the amount of convolution and thus need to be evaluated differently in later 
subsection.

We could see that in the best case, the optimization strategies reduce the amount of 
convolution by \textbf{98\%} (in the case of GtsCe).

%TODO: factors of the network node-pair? path amount and length?

\begin{itemize}
    \item While the amount of path in those states can explain the amount of EC reduction, to get to 
        the amount of convolution, we also need to consider the length of each of those path
    \item By considering the both amount of path and its length to counting the convolution, we could 
        compare consolidation and memoization strategy to the unoptimized version
    \item 
\end{itemize}

@@

@@

\section{Related Works}
\begin{itemize}
    \item Control plane and data plane verifier
    \item Probabilistic Verifier 
    \item Quantitative Verifier
\end{itemize}
